Bird beak's elaboration and exploration

### 2021-05-27 Analysis/03-elaboration_exploration_analyses.Rmd
Illustration of proj/rej

```{r, eval = FALSE, echo = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
```

@@

```{r, eval = FALSE}
set.seed(42)
## Selecting 1000 random covariance matrices from the MCMCglmm
covar_matrices <- get.covar(combined_results, n = 100)
```

We will use these covariance matrices as the basis of the phylogenetic main axes of variation.
These axes correspond to the main axes of variation in the $N_DIMENSIONS$ trait space at different levels: 
 * The whole bird phylogeny level ("phylogeny level");
 * And the named node levels ("clade levels").

@@

### Details for getting the major axes from the variance co-variance matrices

@@

The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `beer::get.axis` ([specific procedure on lines 40-73](https://github.com/TGuillerme/elaboration_exploration_bird_beaks/blob/master/beer/R/get.axes.R)).

>>>>

```{r, eval = TRUE, echo = FALSE}
library(beer)
## Loading the MCMCglmm
load("../Data/processed/combined_results.rda")
## Loading the tree and data
data(demo_data)
data(demo_tree)
## Setting the colour vector
colour_vector <- c("orange", "blue", "darkgreen")
```

@@

```{r, eval = TRUE}
set.seed(42)
## Selecting 1000 random covariance matrices from the MCMCglmm
covar_matrices <- get.covar(combined_results, n = 100)
```

We will use these covariance matrices as the basis of the phylogenetic main axes of variation.
These axes correspond to the main axes of variation in the $N_DIMENSIONS$ trait space at different levels:

 * The whole bird phylogeny level ("phylogeny level");
 * And the named node levels ("clade levels").

@@

## Major axes calculation

@@

> The detailed procedure was adapted from [李哲源's post on Stack Overflow](https://stackoverflow.com/questions/40300217/obtain-vertices-of-the-ellipse-on-an-ellipse-covariance-plot-created-by-care/40316331#40316331) and implemented in `beer::get.axis` ([specific procedure on lines 40-73](https://github.com/TGuillerme/elaboration_exploration_bird_beaks/blob/master/beer/R/get.axes.R)).

```{r}
## Calculating the major axes
major_axes <- get.axes(covar_matrices, centre = "intercept")
```

One can visualise these ellipses and major axes in 2D using the `beer::plot.axes` and `beer::plot.ellipses` functions (based on @ellipse):

```{r}
## Plotting the ellipses
plot.ellipses(covar_matrices, col = c("grey",colour_vector),
              add = FALSE)
plot.axes(major_axes, col = c("grey",colour_vector),
              add = TRUE)
legend("topleft", legend = c("all_clades", levels(demo_data$clade)), col = c("grey", colour_vector), pch = 19)
```

We can then compare these axes either between each other or relative to the tips positions in the trait space using elaboration and exploration scores:

## Elaboration and exploration scores calculations

From these major axes, we can use linear algebra in $n$ dimensions to estimate clades or tips' exploration and elaboration scores.
These scores are a direct interpretation of @endler2005's elaboration and innovation (here exploration) patterns (Fig 1. in @endler2005):
Given a major axis of evolutionary "trajectory" (in the sense of evolutionary trend, not an evolutionary direction), on can calculate the elaboration score as the projection value of any vector defined by a clade or a tip on the major axis and the exploration (or "innovation" in @endler2005) as that vector's rejection.

```{r, echo = FALSE, eval = TRUE}
source("../Functions/test.rotation.R")
```
Figure: in a space with 5 elements: A, B, C, D, E (in grey); where D and E represent the major axes (e.g. the major phylogenetic effect axes from the MCMCglmm), we can rotate and rescale each element so that D and E become the unit vector of length 1 (the black letters D and E) and get the exploration elaboration scores for either the elements (e.g. element A's projection in blue and its rejection in orange) or from any other axes (e.g. the major phylogenetic effect axes from the MCMCglmm from a specific clade).

To measure the projection and rejection scores, we used the following algorithm:

 * Rotate
 * Rescale
 * Project/reject
